\documentclass{article}

\usepackage{../cmsc491}
\fancyhead[R]{CMSC 491, Project Proposal}

\pagestyle{fancy}

\linespread{1.3}

\begin{document}

\begin{center}
    \LARGE Proposal: Occupancy and ``Clutter'' Detection \\
    for Home Automation
\end{center}

Computer vision can be used to leverage sensing capabilities of consumer cameras
and computer hardware for home automation purposes. Home automation is a rapidly
growing field involving applied computer science; it currently suffers from
an abundance of ``output'' hardware, such as smart thermostats and
network-connected light bulbs, but a small number of consumer-accessible
``input'' hardware for sensing home conditions, such as occupancy, motion, and
ambient lighting.\footnote{Such sensors exist, but often involve integrated
circuitry and custom software components.}

Network- or PC-connected visible-light cameras are common in many households,
and hence may allow us to leverage computer vision for numerous sensing tasks,
including occupancy detection. I propose to study both occupancy detection and
``clutter.'' Occupancy detection is an important problem in home automation
applications. Clutter is a measure the number of miscellaneous object objects
in a living space, and could be used to indicate when cleaning up is necessary.
Both applications will be use images taken on consumer cameras (Raspberry Pi and
security cameras) and run on consumer computing hardware suitable for actual
home use.

Occupancy is a related but distinct problem from detecting motion, which is
often solved by passive infrared sensors (PIRs).\footnote{Adafruit PIR kits
    for the Raspberry Pi: \url{https://adafruit.com/products/189}.}
Occupants of a living area may be stationary, and PIRs may be triggered by
non-humanoid motion, such as by pets. Computer vision may be able to solve this
problem elegantly. The popular open source library
OpenCV\footnote{\url{http://opencv.org}}
includes a pre-trained histogram of oriented gradients (HoG) detector which I
expect to use.

Clutter detection appears to be a novel application for computer vision. My
first approach is to use an HoG descriptor on a training set comprised of images
of a house in various degrees of clutter. Another possible method is to analyze
the two-dimensional Fourier transform space of the images: higher frequencies of
edges will indicate many small objects and lower frequencies will reflect
larger, permanent objects in the living space. This problem will likely require
the greatest portion of research time.

Currently, my languages of choice for the project are Python and
Julia,\footnote{\url{http://julialang.org}} a fast scientific programming
language. They have excellent cross-language calling support, and seem
well-suited to the application. I am also aiming to integrate my software with
Idiotic, an open source home automation
framework.\footnote{\url{https://github.com/idiotic/idiotic}} Using Python will
make it very easy to include as a module, and hence release under an open source
license for wider use.

\end{document}
